Revisiting the impact of deepfake technology towards viewer's political choice: A replication of Soubhik Barari (2021)

### 2021-03-18 outputs/paper/paper.Rmd
Update citation & formatting

```{r postfavor, fig.cap="Distribution of feelings towards Sanders in six different situations",echo=FALSE}
#histogram for each treat
EDA<-EDA%>%
  drop_na(post_favor_Warren)

@@

Feature importance plays a vital role in predictive modelling projects. It can provide insights on data, models, and how to reduce dimensionality and select features, thereby improving efficiency and effectiveness of the model prediction. [@guyon2003introduction]

Random Forest is the chosen method for our feature selection model because it can handle multi-dimensional data and output feature importance according to maximized information gain in node splitting. [@nguyen2013random]The random forest algorithm forms a series of classification methods that rely on a combination of several decision trees. According  to the guideline from Liaw, the random forest algorithm is optimized via three parameters in our project: we set ntree=500 (the number of trees), mtry=all the variables in FS dataset divided by 3 (the number of predictors randomly tested at each node) and node size=5 (the minimal size of the terminal node). [@liaw2018classification]

@@

```
Next important variable comes to `sexism`. Lovenduski mentioned that “Women remain significantly under-represented in political life.” A big reason behind this is that people still associate toughness and leadership qualities with men. “Female marginalisation is hardwired into the traditional institutions within which politics takes place.” Of course, things have two sides. American feminists also let Warren have a large number of loyal female supporters. [@lovenduski2014institutionalisation]

@@

We have made a choice between linear and non-linear models. Random forest regression is an alternative form of regression. It works well with both categorical and continuous features. And it can handle nonlinear elements in the model. Normalizing of data is not required as it uses a rule-based approach. However, despite these advantages, a random forest algorithm also has some drawbacks. Although it achieves higher prediction accuracy than linear regression models, random forest regression suffers from weak interpretability. It does not output estimated variable coefficients or p-values to determine significance of each variable, which are indispensable for statistical inference and model interpretation. Besides, it demands more computational power as it builds numerous trees to combine their outputs. Furthermore, the random forest regression model is susceptible to overfitting and the ultimate model performance relies on manual hyperparameter tuning. Consequently, our group decided to choose the multiple linear regression model. [@strobl2008danger]

We are using RStudio to run the multiple linear regression model. The following reasons explain why we decide to choose multiple linear regression[@WinNT]:

@@

R-squared  is a measure of the goodness of fit of the model. A larger R-squared  indicates a closer fit of the model to the data. It is used as an optimality criterion in parameter selection and model selection.[@johansen1980welch]

@@

P-value is used to describe the occurrence possibility of the extreme outcome when the null hypothesis is true. If the p-value is small, it means that the probability of occurrence of the null hypothesis to be true is very small. And if it does occur, we have a reason to reject the null hypothesis. In short, the smaller the p-value, the more significant the result. Usually the threshold for significant p value is set to 0.05.[@ioannidis2018proposal]

@@

The sign of the regression coefficient describes whether there is a positive or negative correlation between each feature variable and the dependent variable (Warren affective score). A positive coefficient means that as the value of the independent variable increases, the average value of the affective score also tends to increase. A negative coefficient indicates that as the independent variable increases, the affective score tends to decrease.[@uyanik2013study]

@@

```{r diagnostic, fig.cap = "Regression diagnostics", echo=FALSE, warning=FALSE, fig.width=10, fig.height=10}
par(mfrow = c(2, 2))
plot(model)
```

```{r model_improved, echo=FALSE, warning=FALSE, fig.width=10, fig.height=2}
#model2 <- lm(sqrt(post_favor_Warren)~., data = FS)
#autoplot(model2,3)
```
**Residual and fitted plot: **
This plot  is used to check the linear relationship assumptions. A horizontal line, without distinct patterns is an indication for a linear relationship. Here we see that linearity seems to hold reasonably well, as the red line is close to the horizontal level (at 0). This suggests that we can assume a linear relationship between the predictors and the outcome variables. [@bruce2020practical]

**Normal QQ plot:  **
It is used to examine whether the residuals are normally distributed. It’s good if residuals points follow the straight dashed line. Here we can see our data points closely follow the dash line, we can assume normality.[@bruce2020practical]

**Scale Location plot:**
It is used to check the homogeneity of variance. A horizontal line with equally spread points is a good indication of homoscedasticity. however, in our case, we can see the blue line has a concave pattern, which indicates our model’s homoscedasticity is not good enough. [@bruce2020practical]

@@

**Residual vs leverage plot:**
It is used to identify influential cases, that is extreme values that might influence the regression results when included or excluded from the analysis. In our case, the fitness is not good, we can find points with high leverage influence.[@bruce2020practical]

@@

* Gender: Female
*	Education Level: lower than high school degree 
* Income: Less than $25K
*	Ethnicity: Asian
* Region: Midwest
*	Response_wave_ID: SV_OxlqWIOfO10wuYI
* Device: Desktop
*	Age group: Less than 65 years old
*	Partisan: Democrat
* Media condition: Control
*	Prompt: Control

@@

### Internal validity
The internal validity of the model can be justified by the procedures of sampling, randomization in experiment intervention, controlling for confounders and having sufficient sample size in the regression model, so that alternative explanations but the treatment can be eliminated from the cause of effect on the outcome.

@@

### External validity
External validity of conclusion on deepfakes was enhanced by the use of 5 customized deepfake contents instead of recycling one well-recognized deepfake example in the treatment experiment. The sampling population is from the U.S. national level, so that the study conclusion and subject trait summary are applicable to the domestic level. The study consists of two waves between Sep. 29th and Oct. 29th. The repetition of experiment conducted at different time periods is to smooth out some possible time effects in the result. However, the model result cannot be generalized to outside the U.S. due to the limitation in the topic itself - U.S. presidential election.

@@

### Opportunities
The study investigates the impact of deepfake media on political misinformation through two field experiments conducted within a month period through the topic of 2020 Democratic presidential election. The external validity of the study can be further improved in order for the result to be generalizable to other political parties, time periods and situations. 
**Time window**: for the experiment time window, it can be prolonged or randomized in a series of experiment repetitions as long as they are conducted after the time point when the in-party presidential nominee has been finalized.

@@

## Differences and difficulties
### Differences
**Feature selection**: Before building the multiple linear regression model, feature engineering was used to visualize feature importance and select the top 20 most important features from a total of 100 variables with the Random Forest classifier [@citeRandomforest]. The tree-based algorithm splits nodes based on how well they improve the purity of the node through maximizing information gain. The most informative features are at the start of the trees, and less relevant features are places at lower nodes. This process naturally ranks the importance of features, therefore it can be used for feature selection.

>>>>

```{r postfavor, fig.cap="Distribution of feelings towards Sanders in six different situations",echo=FALSE, fig.width=6, fig.height=4}
#histogram for each treat
EDA<-EDA%>%
  drop_na(post_favor_Warren)

@@

Feature importance plays a vital role in predictive modelling projects. It can provide insights on data, models, and how to reduce dimensionality and select features, thereby improving efficiency and effectiveness of the model prediction [@guyon2003introduction].

Random Forest is the chosen method for our feature selection model because it can handle multi-dimensional data and output feature importance according to maximized information gain in node splitting [@nguyen2013random]. Random forest algorithm forms a series of classification methods that rely on a combination of several decision trees. According  to the guideline from Liaw [@nguyen2013random], the random forest algorithm is optimized via three parameters in our project: we set ntree=500 (the number of trees), mtry=all the variables in FS dataset divided by 3 (the number of predictors randomly tested at each node) and node size=5 (the minimal size of the terminal node).

@@

```
Next important variable comes to `sexism`. Lovenduski [@lovenduski2014institutionalisation] mentioned that “Women remain significantly under-represented in political life.” A big reason behind this is that people still associate toughness and leadership qualities with men. “Female marginalisation is hardwired into the traditional institutions within which politics takes place.” Of course, things have two sides. American feminists also let Warren have a large number of loyal female supporters.

@@

We have made a choice between linear and non-linear models. Random forest regression is an alternative form of regression. It works well with both categorical and continuous features. And it can handle nonlinear elements in the model. Normalizing of data is not required as it uses a rule-based approach. However, despite these advantages, a random forest algorithm also has some drawbacks [@strobl2008danger]. Although it achieves higher prediction accuracy than linear regression models, random forest regression suffers from weak interpretability. It does not output estimated variable coefficients or p-values to determine significance of each variable, which are indispensable for statistical inference and model interpretation. Besides, it demands more computational power as it builds numerous trees to combine their outputs. Furthermore, the random forest regression model is susceptible to overfitting and the ultimate model performance relies on manual hyperparameter tuning. Consequently, our group decided to choose the multiple linear regression model.

We are using RStudio to run the multiple linear regression model. The following reasons explain why we decide to choose multiple linear regression [@WinNT]:

@@

R-squared  is a measure of the goodness of fit of the model. A larger R-squared  indicates a closer fit of the model to the data. It is used as an optimality criterion in parameter selection and model selection [@johansen1980welch].

@@

P-value is used to describe the occurrence possibility of the extreme outcome when the null hypothesis is true [@ioannidis2018proposal]. If the p-value is small, it means that the probability of occurrence of the null hypothesis to be true is very small. And if it does occur, we have a reason to reject the null hypothesis. In short, the smaller the p-value, the more significant the result. Usually the threshold for significant p value is set to 0.05.

@@

The sign of the regression coefficient describes whether there is a positive or negative correlation between each feature variable and the dependent variable (Warren affective score). A positive coefficient means that as the value of the independent variable increases, the average value of the affective score also tends to increase. A negative coefficient indicates that as the independent variable increases, the affective score tends to decrease [@uyanik2013study].

@@

```{r diagnostic, fig.cap = "Regression diagnostics", echo=FALSE, warning=FALSE, fig.width=8, fig.height=8}
par(mfrow = c(2, 2))
plot(model)
```

```{r model_improved, echo=FALSE, warning=FALSE, fig.width=10, fig.height=2}
#model2 <- lm(sqrt(post_favor_Warren)~., data = FS)
#autoplot(model2,3)
```
**Residual and fitted plot: **
This plot  is used to check the linear relationship assumptions. A horizontal line, without distinct patterns is an indication for a linear relationship. Here we see that linearity seems to hold reasonably well, as the red line is close to the horizontal level (at 0). This suggests that we can assume a linear relationship between the predictors and the outcome variables [@bruce2020practical].

**Normal QQ plot:  **
It is used to examine whether the residuals are normally distributed. It’s good if residuals points follow the straight dashed line. Here we can see our data points closely follow the dash line, we can assume normality [@bruce2020practical].

**Scale Location plot:**
It is used to check the homogeneity of variance. A horizontal line with equally spread points is a good indication of homoscedasticity. however, in our case, we can see the blue line has a concave pattern, which indicates our model’s homoscedasticity is not good enough [@bruce2020practical].

@@

**Residual vs leverage plot:**
It is used to identify influential cases, that is extreme values that might influence the regression results when included or excluded from the analysis. In our case, the fitness is not good, we can find points with high leverage influence [@bruce2020practical].

@@

* `Gender`: Female
*	`Education Level`: Lower than high school degree 
* `Income`: Less than $25K
*	`Ethnicity`: Asian
* `Region`: Midwest
*	`Response_wave_ID`: SV_OxlqWIOfO10wuYI
* `Device`: Desktop
*	`Age group`: Younger than 65 years old
*	`Partisan`: Democrat
* `Media condition`: Control
*	`Promp`t: Control

@@

### Internal validity
The internal validity of the model can be justified by the procedures of sampling, randomization in experiment intervention, controlling for confounders and having sufficient sample size in the regression model [@citeValidity], so that alternative explanations but the treatment can be eliminated from the cause of effect on the outcome.

@@

### External validity
External validity of conclusion on deepfakes was enhanced by the use of 5 customized deepfake contents instead of recycling one well-recognized deepfake example in the treatment experiment [@externalv]. The sampling population is from the U.S. national level, so that the study conclusion and subject trait summary are applicable to the domestic level. The study consists of two waves between Sep. 29th and Oct. 29th. The repetition of experiment conducted at different time periods is to smooth out some possible time effects in the result. However, the model result cannot be generalized to outside the U.S. due to the limitation in the topic itself - U.S. presidential election.

@@

### Opportunities

The study investigates the impact of deepfake media on political misinformation through two field experiments conducted within a month period through the topic of 2020 Democratic presidential election. The external validity of the study can be further improved in order for the result to be generalizable to other political parties, time periods and situations.

**Time window**: for the experiment time window, it can be prolonged or randomized in a series of experiment repetitions as long as they are conducted after the time point when the in-party presidential nominee has been finalized.

@@

## Differences and difficulties

### Differences

**Feature selection**: Before building the multiple linear regression model, feature engineering was used to visualize feature importance and select the top 20 most important features from a total of 100 variables with the Random Forest classifier [@citeRandomforest]. The tree-based algorithm splits nodes based on how well they improve the purity of the node through maximizing information gain. The most informative features are at the start of the trees, and less relevant features are places at lower nodes. This process naturally ranks the importance of features, therefore it can be used for feature selection.
